A Parallel General Purpose Multi-Objective Optimization Framework, with Application To Electron Beam Dynamics

### 2018-12-21 paper/submission/ZW10157-response-1.tex
updating bib

{\bf Our response:} {\color{blue} The code sections have been removed.}

@@

@@

Section VI

Code and input file details should be in Supplemental Material

Section V

subsection B

Code and input file details should be in Supplemental Material

{\bf Our response:} {\color{blue} The code has been removed from the updated manuscript. Interested parties can refer to the repository.}

@@

{\bf Our response:} {\color{blue} This table shows input parameters for one optimized solution (the solution in Fig. 12). 
The table title has been changed to the following: "Input Parameters for the optimized solution shown in Figures~11 and 12."}

>>>>

{\bf Our response:} {\color{blue} The code examples in sections IV-V have been removed.}

@@

{\bf Our response:} {\color{blue} We agree that some references of import were missing and should have been included.
Especially, the following:

}

@@

Section VI: 
Code and input file details should be in Supplemental Material

Section V, subsection B:
Code and input file details should be in Supplemental Material

{\bf Our response:} {\color{blue} The code examples in sections VI-V has been removed from the updated manuscript. 
	A sentence and link to the repository have been provided. Interested parties can refer to the repository.}

@@

{\bf Our response:} {\color{blue} This table shows input parameters for one optimized solution (the solution in Fig. 12). 
The table title has been changed to the following: ``Input Parameters for the optimized solution shown in Figures~11 and 12."}

### 2018-12-21 paper/submission/prab.tex
removing grids from plots in figure 10
updating bib

A modular multi-objective software framework was developed (see
 Fig.~\ref{fig:framenetwork}) where the core functionality of the optimizer is decoupled from
 the ``beam dynamics'' but fully integrated in the OPAL \cite{opal} framework. 
To that end, we use a master/slave mechanism where a master process governs a
 set of slave processes given some computational tasks (beam dynamics simulation) to complete.
This separation allows easy interchange of optimization algorithms, forward
  solvers and optimization problems.
A ``pilot'' coordinates all efforts between the optimization algorithm and the
  beam dynamics task. In the sequel we will also use the notion of ``forward solver'' to indicate the beam dynamics task.
This forms a robust and general framework for massively parallel
  multi-objective optimization.
Currently the framework offers one concrete optimization algorithm, an
  evolutionary algorithm employing a \textsc{NSGA-II} selector \cite{pisa}.
Normally, simulation based approaches are plagued by the trade-of between
  level of detail and time to solution.
This problem is addressed later in Section~(\ref{awa:subsection:test}) by using forward solvers with different time and
  resolution complexity.

@@

In general, it is not possible to get the maximal performance for the lowest
  price and a trade-off decision between performance and price has to be
  reached (see Fig.~\ref{fig:tradeoff}).
Since not every choice is equally profitable for the buyer (for example, car
  $x_4$ costs as much as $x_2^*$ but offers less performance), we pick
  trade-offs (red points) that are essentially ``equally optimal'' in both
  conflicting objectives, meaning, we cannot improve one point without hurting
  at least one other solution.
This is known as \textit{Pareto optimality}.
The set of Pareto optimal points (blue curve) forms the Pareto front or
  surface.
All points on this surface are considered to be Pareto optimal.

Once the shape of the Pareto front has been determined, the buyer can
  \textit{specify preference}, balancing the features by observing the
  effect on the optimality criteria, converging to the preferred solution.
This is called \textit{a posteriori} preference specification since 
  a solution is selected after all possible trade-offs have been presented to us.
An alternative is to specify preference \textit{a priori}, e.g. by
  weighting (specifying preference before solving the problem) and combining
  all objectives into one and applying a single-objective method to solve the
  problem (yielding only one solution).
In many situations preference is not known \textit{a priori} and an
  \textit{a posteriori} preference specification helps conveying a deeper
  understanding of the solution space.
The Pareto front can be explored and the impact of a trade-off decision then
  becomes visible.

@@

The framework contains a wrapper implementing the API mentioned in
  Listing 3 for using \textsc{OPAL} \cite{opal} as the forward solver.
\textsc{OPAL} provides different trackers for cyclotrons and linear
  accelerators with satisfactory parallel performance \cite{akir:09}.

@@

%
Using the default expression primitives this can be stated in an input file
  as:
%
%\vspace{0.2cm}

\begin{samepage}
	\begin{flushleft}
		\begin{Verbatim}[fontsize=\scriptsize]
		d1: DVAR, VARIABLE="x1", LOWERBOUND="-1.0", UPPERBOUND="1.0";
		d2: DVAR, VARIABLE="x2", LOWERBOUND="-1.0", UPPERBOUND="1.0";
		d3: DVAR, VARIABLE="x3", LOWERBOUND="-1.0", UPPERBOUND="1.0";

@@

		objs:    OBJECTIVES = (obj1, obj2);
		dvars:   DVARS = (d1, d2, d3);
		constrs: CONSTRAINTS = ();
		opt: OPTIMIZE, OBJECTIVES=objs, DVARS=dvars, CONSTRAINTS=constrs;
		\end{Verbatim}
	\end{flushleft}
\end{samepage}

>>>>

A modular multi-objective software framework was developed (see
 Fig.~\ref{fig:framenetwork}) where the core functionality of the optimizer is decoupled from
 the ``beam dynamics'' but fully integrated in the OPAL \cite{opal} framework. 
To that end, we use a master/slave mechanism where a master process governs a
 set of slave processes given some computational tasks (beam dynamics simulation) to complete.
This separation allows easy interchange of optimization algorithms, forward
  solvers and optimization problems.
A ``pilot'' coordinates all efforts between the optimization algorithm and the
  beam dynamics task. In the sequel we will also use the notion of ``forward solver'' to indicate the beam dynamics task.
This forms a robust and general framework for massively parallel
  multi-objective optimization.
Currently the framework offers one concrete optimization algorithm, an
  evolutionary algorithm employing a \textsc{NSGA-II} selector \cite{pisa}.
Normally, simulation based approaches are plagued by the trade off between
  level of detail and time to solution.
This problem is addressed later in Section~(\ref{awa:subsection:test}) by using forward solvers with different time and
  resolution complexity.

@@

In general, it is not possible to get the maximal performance for the lowest
  price and a trade off decision between performance and price has to be
  reached (see Fig.~\ref{fig:tradeoff}).
Since not every choice is equally profitable for the buyer (for example, car
  $x_4$ costs as much as $x_2^*$ but offers less performance), we pick
  trade offs (red points) that are essentially ``equally optimal'' in both
  conflicting objectives, meaning, we cannot improve one point without hurting
  at least one other solution.
This is known as \textit{Pareto optimality}.
The set of Pareto optimal points (blue curve) forms the Pareto front or
  surface.
All points on this surface are considered to be Pareto optimal.

Once the shape of the Pareto front has been determined, the buyer can
  \textit{specify preference}, balancing the features by observing the
  effect on the optimality criteria, converging to the preferred solution.
This is called \textit{a posteriori} preference specification since 
  a solution is selected after all possible trade offs have been presented to us.
An alternative is to specify preference \textit{a priori}, e.g. by
  weighting (specifying preference before solving the problem) and combining
  all objectives into one and applying a single-objective method to solve the
  problem (yielding only one solution).
In many situations preference is not known \textit{a priori} and an
  \textit{a posteriori} preference specification helps conveying a deeper
  understanding of the solution space.
The Pareto front can be explored and the impact of a trade off decision then
  becomes visible.

@@

The framework contains a wrapper implementing the API mentioned in
  Listing 3 for using \textsc{OPAL} \cite{opal} as the forward solver.
\textsc{OPAL} provides different trackers for cyclotrons and linear
  accelerators with satisfactory parallel performance. % \cite{akir:09}.

@@

%
Using the default expression primitives this can be stated in an input file
  as:
%
%\vspace{0.2cm}
\iffalse
\begin{samepage}
	\begin{flushleft}
		\begin{Verbatim}[fontsize=\scriptsize]
		d1: DVAR, VARIABLE="x1", LOWERBOUND="-1.0", UPPERBOUND="1.0";
		d2: DVAR, VARIABLE="x2", LOWERBOUND="-1.0", UPPERBOUND="1.0";
		d3: DVAR, VARIABLE="x3", LOWERBOUND="-1.0", UPPERBOUND="1.0";

@@

		objs:    OBJECTIVES = (obj1, obj2);
		dvars:   DVARS = (d1, d2, d3);
		constrs: CONSTRAINTS = ();
		opt: OPTIMIZE, OBJECTIVES=objs, DVARS=dvars, CONSTRAINTS=constrs;
		\end{Verbatim}
	\end{flushleft}
\end{samepage}
\fi

### 2018-12-21 paper/submission/theor-experiment.tex
fixing figure names in paper
updating bib

\subsubsection{Hyper parameter Scan}
While the optimization problem and goals were well defined (Subsection \ref{awaproblem}), 
it was not clear what the best hyper parameters for the genetic algorithm would be.
These parameters include gene mutation probability, mutation probability, 
recombination probability, number of individuals, 
and number of generations to complete. 
Given the beam line in Fig.~\ref{awa-linac},
four small optimization experiments were done with various hyper parameters. 
Similar to the time step scan, 
the goal of this exercise was to determine which set of optimization
parameters strongly influence the results, 
and whether there was a time to solution difference.
From here on, we will reference each experiment as ex-1, ex-2, ex-3, and ex-4
as shown in Table \ref{extable}. 
\begin{table}%[h!]
	\begin{center}
		\caption{Input Parameters for initial twenty four hour AWA optimization experiments. 
			The gene mutation probability was equal to the mutation probability (not shown) in all four experiments. 
			The max number of individuals per generation was~80.}
		\label{extable}
		\begin{tabular*}{0.48\textwidth}{l @{\extracolsep{\fill}} C C D }
			\hline\noalign{\smallskip}
			& Gene Mutation Probability & Recombination Probability & Number of completed generations \\
			\noalign{\smallskip}\hline\noalign{\smallskip}
			ex-1 &  0.1  & 0.9  &  96 \\
			ex-2 &  0.3  & 0.7  &  81 \\
			ex-3 &  0.8  & 0.2  &  53 \\
			ex-4 &  0.01 & 0.09 &  95 \\
			\noalign{\smallskip}\hline
		\end{tabular*}
	\end{center}
\end{table}

@@

The objectives include: $\varepsilon_{x}\left(s = s_1\right)\text{, } \varepsilon_{x}\left(s = s_2\right)$, $\text{rms}_{s}\left(s = s_1\right)\text{, and }  \text{rms}_{s}\left(s = s_2\right)$. 
The OPAL input file for ex-3 is given as an example to show how optimization and design variables are defined:

@@

\subsubsection{TBA Optimization Problem}
With computational and hyper parameters set, 
the full optimization problem of interest is explored.
The objectives (beam sizes and energy spread) are calculated at 
$s_3=19.4$~m, located downstream of the septum, see Fig.~\ref{awa-tba}. 
Given the longitudinal location of $s_3$ (unless otherwise noted), 
we define the objectives and input parameters as:
\begin{align}
\text{min}  \quad & \text{rms}_{x}, \quad \text{rms}_{y} \label{eq:awa:p1}\\
& \text{rms}_{px}, \quad \text{rms}_{py}, \label{eq:awa:p2}\\
& \text{rms}_{s}, \quad dE \label{eq:awa:p4} \\
\text{constraints} \quad & rms_x < 0.1 \, (m) |_{s=s_1}\label{eq:awa:c1}\\
\quad & rms_y < 0.1\, (m) |_{s=s_1}\, \label{eq:awa:c2}\\
\quad & |rms_y - rms_x | < 0.005 \, (m) |_{s=s_1}\label{eq:awa:c3}\\
\text{subject to} \quad & q = 40 \, (nC) \label{eq:awa:firstconstr}\\
\quad & \text{Volt}_{\text{Gun}} = 64 \, \, (\text{MV/m}) \label{eq:awa:lastconstr}\\
\quad & \text{Volt}_{\text{Linac}} = 24 \, \text{or} \, 25\,\, (\text{MV/m}) \\
\quad & R_x = R_y = 9 \, (mm) \label{eq:awa:firstdvar}\\
\quad & \phi_{\text{gun}} =-20^\circ \label{eq:awa:gphidvar}\\
\quad & \phi_{\text{linac}} =-20^\circ \label{eq:awa:lastdvar}
\end{align}

@@

With the time steps and hyper parameters set by the work in Section \ref{awa:subsection:test}, 
the full optimization problem described in \ref{awaproblem} was run for 200 generations.
The initial number of individuals was fixed at 656, 
and the minimum number individuals in later generations was fixed at 328. 
These numbers were in part based on the architecture of the KNL's. 
Since each simulation takes 8 cores, and there are 64 cores per KNL node, 
a large population size that would fit evenly on these resources was chosen. 
Again, the location of optimization is $s_3=19.4$ [m]. 
\begin{figure}
	\begin{center}		
		\begin{tikzpicture}[scale=1]
		\node (fig1) at (0,0)
		{\includegraphics[width=0.5\textwidth]{figures/xy_vs_pxy_pareto_front_quads_before_Q5}};
		\node[star, star points=5, star point ratio=0.5, fill=yellow, draw](key) at (3.35,-1.7) {};
		\end{tikzpicture}
	\end{center} 
		\caption{Pareto front comparing transverse beam sizes ($\sigma_{x,y}$) and transverse momentum ($\sigma_{px,py}$). The yellow star indicates the 
		point plotted in Fig.~\ref{fig:stat}}
	\label{fig:pareto1}
\end{figure}

@@

\begin{figure}
	\includegraphics[width=0.48\textwidth]{figures/xy-max-min-sigma-paper}
	\caption{Optimized beam sizes along high charge beam line. The gun is located at $s=0$, 
	both x and y beam sizes are shown. The black line represents the relevant beam line aperture, while
	the green line indicates the location of the optimization.}
	\label{fig:stat}
\end{figure}
\begin{table}%[h!]
	\begin{center}
		\caption{Input Parameters for large scale TBA optimization runs.}
		\label{tab:designopt}   
		\begin{tabular}{l|c|c}
			\noalign{\smallskip}\hline\noalign{\smallskip}
			%\rowcolor{blue!30} 
			\textbf{Design Variable} & \textbf{Unit}	&  \textbf{Value}  \\ 
			\noalign{\smallskip}\hline\noalign{\smallskip}
			{Buck Focusing Solenoid} & amps	& 478 \\
			Matching Solenoid &	amps	& 197	  \\
			Quadrupole 1& T-m		& -0.8	\\ 
			Quadrupole 2& T-m		& 0.9	\\
			Quadrupole 3 & T-m		& 0.8	\\
			Quadrupole 4 & T-m		& -1.0	\\ 
			Bunch Length & mm 		& 1.5	\\
			\noalign{\smallskip}\hline\noalign{\smallskip}
		\end{tabular}
	\end{center}
\end{table}

>>>>

\subsubsection{Hyper parameter Scan}
While the optimization problem and goals were well defined (Subsection \ref{awaproblem}), 
it was not clear what the best hyper parameters for the genetic algorithm would be.
These parameters include gene mutation probability, mutation probability, 
recombination probability, number of individuals, 
and number of generations to complete. 
Given the beam line in Fig.~\ref{awa-linac},
four small optimization experiments were done with various hyper parameters. 
Similar to the time step scan, 
the goal of this exercise was to determine which set of optimization
parameters strongly influence the results, 
and whether there was a time to solution difference.
From here on, we will reference each experiment as ex-1, ex-2, ex-3, and ex-4
as shown in Table \ref{extable}. 
\begin{table}%[h!]
	\begin{center}
		\caption{Input Parameters for initial twenty four hour AWA optimization experiments. 
			The gene mutation probability was equal to the mutation probability (not shown) in all four experiments. 
			The max number of individuals per generation was~80.}
		\label{extable}
		\begin{tabular*}{0.48\textwidth}{l @{\extracolsep{\fill}}| C | C | D }
			\hline\noalign{\smallskip}
			& Gene Mutation Probability & Recombination Probability & Number of completed generations \\
			\noalign{\smallskip}\hline\noalign{\smallskip}
			ex-1 &  0.1  & 0.9  &  96 \\
			ex-2 &  0.3  & 0.7  &  81 \\
			ex-3 &  0.8  & 0.2  &  53 \\
			ex-4 &  0.01 & 0.09 &  95 \\
			\noalign{\smallskip}\hline
		\end{tabular*}
	\end{center}
\end{table}

@@

The objectives include: $\varepsilon_{x}\left(s = s_1\right)\text{, } \varepsilon_{x}\left(s = s_2\right)$, $\text{rms}_{s}\left(s = s_1\right)\text{, and }  \text{rms}_{s}\left(s = s_2\right)$. 
The OPAL input file for ex-3, and subsequent optimization runs in this paper, can be found at in following repository: 
\url{https://github.com/nneveu/awa-tba}.
%is given as an example to show how optimization and design variables are defined:

@@

\subsubsection{TBA Optimization Problem}
With computational and hyper parameters set, 
the full optimization problem of interest is explored.
The objectives (beam sizes and energy spread) are calculated at 
$s_3=19.4$~m, located downstream of the septum, see Fig.~\ref{awa-tba}. 
Given the longitudinal location of $s_3$ (unless otherwise noted), 
we define the objectives and input parameters as:
\begin{align}
\text{min}  \quad & \text{rms}_{x}, \quad \text{rms}_{y} \label{eq:awa:p1}\\
& \text{rms}_{px}, \quad \text{rms}_{py}, \label{eq:awa:p2}\\
& \text{rms}_{s}, \quad \text{dE} \label{eq:awa:p4} \\
\text{constraints} \quad & \text{rms}_x < 0.1 \, (\text{m}) |_{s=s_1}\label{eq:awa:c1}\\
\quad & \text{rms}_y < 0.1\, (m) |_{s=s_1}\, \label{eq:awa:c2}\\
\quad & |\text{rms}_y - \text{rms}_x | < 0.005 \, (\text{m}) |_{s=s_1}\label{eq:awa:c3}\\
\text{subject to} \quad & q = 40 \, (\text{nC}) \label{eq:awa:firstconstr}\\
\quad & \text{Volt}_{\text{Gun}} = 64 \, \, (\text{MV/m}) \label{eq:awa:lastconstr}\\
\quad & \text{Volt}_{\text{Linac}} = 24 \, \text{or} \, 25\,\, (\text{MV/m}) \\
\quad & R_x = R_y = 9 \, (\text{mm}) \label{eq:awa:firstdvar}\\
\quad & \phi_{\text{gun}} =-20^\circ \label{eq:awa:gphidvar}\\
\quad & \phi_{\text{linac}} =-20^\circ \label{eq:awa:lastdvar}
\end{align}

@@

With the time steps and hyper parameters set by the work in Section \ref{awa:subsection:test}, 
the full optimization problem described in \ref{awaproblem} was run for 200 generations.
The initial number of individuals was fixed at 656, 
and the minimum number individuals in later generations was fixed at 328. 
These numbers were in part based on the architecture of the KNL's. 
Since each simulation takes 8 cores, and there are 64 cores per KNL node, 
a large population size that would fit evenly on these resources was chosen. 
Again, the location of optimization is $s_3=19.4$ [m]. 
\begin{figure}
	\begin{center}		
		\includegraphics[width=0.65\textwidth]{figures/paperedit_pareto_front_Q5_xy_vs_pxpy}
	\end{center} 
		\caption{Pareto front comparing transverse beam sizes (rms$_{x,y}$) and transverse momentum (rms$_{px,py}$). The yellow star indicates the 
		point plotted in Fig.~\ref{fig:stat}}
	\label{fig:pareto1}
\end{figure}

@@

\begin{figure}
	\includegraphics[width=0.65\textwidth]{figures/xy-max-min-sigma-paper}
	\caption{Optimized beam sizes along high charge beam line. The gun is located at $s=0$, 
	both x and y beam sizes are shown. The black line represents the relevant beam line aperture, while
	the green line indicates the location of the optimization.}
	\label{fig:stat}
\end{figure}
\begin{table}%[h!]
	\begin{center}
		\caption{Input Parameters for the optimized solution shown in Figures~\ref{fig:pareto1} and \ref{fig:stat}.}
		\label{tab:designopt}   
		\begin{tabular}{l|c|c}
			\noalign{\smallskip}\hline\noalign{\smallskip}
			%\rowcolor{blue!30} 
			\textbf{Design Variable} & \textbf{Unit}	&  \textbf{Value}  \\ 
			\noalign{\smallskip}\hline\noalign{\smallskip}
			{Buck Focusing Solenoid} & amps	& 478 \\
			Matching Solenoid &	amps	& 197	  \\
			Quadrupole 1& T-m		& -0.8	\\ 
			Quadrupole 2& T-m		& 0.9	\\
			Quadrupole 3 & T-m		& 0.8	\\
			Quadrupole 4 & T-m		& -1.0	\\ 
			Bunch Length & mm 		& 1.5	\\
			\noalign{\smallskip}\hline\noalign{\smallskip}
		\end{tabular}
	\end{center}
\end{table}
